\vspace{-12px}
\section{Outline of the Thesis}
\vspace{-10px}
\label{sec:outlinethesis}
\textbf{Chapter 1} of the thesis contains a general introduction to the research problem of verification C programs against a functional specification.
We take a C program and its analogue in a safe functional language, and contrast their differences. This helps us motivate the problem and its solution.
We finish with our contributions.

In \textbf{Chapter 2}, we constrain the programs being considered by formulating the problem statement. This helps us define the scope of our solution.
We introduce a custom minimal functional language called `Spec' and define the necessary terminology used in the rest of the thesis.

\textbf{Chapter 3} starts with background on program equivalence, bisimulation relation and product program.
The rest of the chapter gradually introduces our first contribution: A Proof Discharge Algorithm and related sub-procedures with the help
of two example programs. We also introduce a program representation of values, called `reconstruction program'.

Next, we formalize previously discussed topics in \textbf{Chapter 4}. We begin with the specification of our custom language `Spec'.
We give a detailed description of (a) a counterexample-guided search algorithm for finding a bisimulation relation and (b) a counterexample-guided
invariant inference procedure. There two procedures with our proof discharge algorithm allow automatic end-to-end equivalence checking.
We formalize handling of procedure calls in `Spec' as well as C, and finish with a dataflow analysis formulation of a pointer analysis
used by our equivalence checker.

\textbf{Chapter 5} introduces a program graph representation of values, called `deconstruction procedure', similar to `reconstruction procedure' as introduced in \textbf{Chapter 3}.
We motivate it by listing its advantages and give an algorithm to convert expressions to this representation.
This helps us simplify our proof discharge algorithm.

In \textbf{Chapter 6}, we introduce our automatic equivalence checker tool named \toolName{}, based on our proof discharge algorithm
and counterexample-guided search procedures.
\toolName{} is evaluated on a large variety of C programs involving lists, strings, trees and matrices.
This includes C programs taken from C library implementations as well as manually written programs. We show that our equivalence checker is able
to prove equivalence of a single specification with multiple of C implementations, each varying in its data layouts and algorithmic
strategies.

Finally, \textbf{Chapter 7} discusses the limitations of our algorithm and compares it with some related work. We note our major ideas and finish with
some potential future improvements to our algorithm.